Approximating the Number of COVID-19 Infections in the United States Across 2021

### 2023-07-24 reports/state_figures.Rmd
add interpretation of table + cross correlations figure

>>>>

Below, we see that although the implementation where $\Pr(S_1|untested)$ was at the survey as well as the implementation did not vary by state or date, the mean simulation interval width was actually lower, indicating the greater concordance was not merely a result of wider bounds.

state_res %>%
  select(-c(date)) %>%
  distinct() %>%
  mutate(interval_length =exp_cases_ub - exp_cases_lb) %>%
  group_by(vlabel) %>%
  summarize(mean_interval_length = mean(interval_length))

### 2023-07-24 writing/manuscript.Rmd
add interpretation of table + cross correlations figure

In Figure \@ref(fig:simulation-intervals-state), we see the simulation intervals for all two-week intervals and all states as well as the 95% Covidestim credible intervals summed to be on the same time scale. 
We see agreement is much higher before the time period spanning December 2021 through January 2022, where Covidestim intervals tend to be higher than the probabilistic bias intervals. This period corresponds to the Omicron wave of the pandemic (Supplementary Figure michigan-variant). This was a major shift in the pandemic: the Omicron variant is highly transmissible compared to previous variants, and it has immune invasion capacity, which means vaccines provides less protection against infection [@andrews2022]. Omicron also is associated with a lower infection fatality ratio [@liu2022], as well as milder infection overall, which may influence testing behavior. Omicron's rise was rapid too, presumably as a result of this enhanced transmissibility and immune escape;  it became the dominant variant over the course of a single month.

@@

@@

We focus on Massachussetts because this state is the only state that reports both positive and total tests and has the most counties where wastewater data was reported throughout the entire time period of interest. Wastewater data here is useful in the sense that it is inherently not subject to the same biases as voluntary testing data, since it captures anyone in the wastewater catchment area, and as such is not influenced by more symptomatic people being more likely to be tested or differences in access to testing.

@@

```{r testrate-v-ratio, fig.cap  = 'The ratio of the median estimated infections to observed infections plotted against the testing rate, where the testing rate is calculated as the total number tested in a two-week interval over the population size. When the priors are the same for all time intervals, there is minimal variability relationship between the testing rate and the ratio of estimated to observed infections, since the correction for incomplete testing and diagnostic test inaccuracy is identical for each time-interval. However, when we allow $\\beta$ or $\\Pr(S_1|\\text{untested})$ to vary over time, there is more variability in the relationship. A horizontal line in red at 1 is included to reference; a ratio of exactly 1 would indicate no infections went unobserved.'}

@@

We see in Figure \@ref(fig:correlations)a) that for all implementations, the probabilistic bias estimates are highly correlated with wastewater concentrations, and in Figure \@ref(fig:correlations)b) we see they also highly correlated with the Covidestim estimates. The correlations in both cases are very similar between implementations, differing by less than 0.01.

@@

Table \@ref(table:coverage) is

@@

county_coverage %>%
  bind_rows(state_coverage) %>%
  group_by(`Geographic Scale`) %>%
  arrange(desc(`Percent Contained in Interval`),
          .by_group=TRUE) %>%
  ungroup() %>%
  select(-`Geographic Scale`) %>%
  kbl(caption = "\\label{table:coverage}Coverage of Covidestim Medians at the County and State Levels",
      booktabs=TRUE,  
      escape=FALSE,
      digits=3) %>%
  kable_styling(latex_options=c("scale_down","HOLD_position"))  %>%
  group_rows(start_row=1, end_row =4, group_label= "County", 
             latex_gap_space = "0.7em",
             hline_after = TRUE) %>%
  group_rows(start_row=5, end_row =8, group_label= "State",
             latex_gap_space = "0.7em",
             hline_after = TRUE) %>%
  row_spec(1:(row_num-1),  hline_after = TRUE, 
           extra_css = "border-bottom: 1px solid;") %>%
  row_spec(0, font_size=11, bold=TRUE) %>%
  footnote(general = paste0("The percent of simulation intervals where the Covidestim median falls below, within, or above the interval,",
  "when considering all simulation intervals for that implementation and geographic scale.",
  "Implementations are ordered from those with the highest proportion of Covidestim medians contained in the interval, to the lowest."),
  general_title ="", footnote_as_chunk=TRUE, threeparttable=TRUE)

@@

In Figure \@ref(fig:cross-corr-wastewater), we see that among all counties except one in Massachsuetts, the maximum correlation between mean wastewater concentration and biweekly infection estimates was obtained at a lag of zero, that is, there  typically there was no lag among the biweekly infection estimates and mean wastewater concentrations. This is to be expected given the lead time for wastewater concentrations, when observed, is less than 2 weeks [@olesen2021a]. In two counties, Hamsphire and Barnstable, the lag between the probabilistic bias counts and wastewater concentration was less than that between the observed positive tests and wastewater concentrations.

```{r cross-corr-wastewater, out.width="70%"}

include_graphics(here("figures/cross_correlation_observed_pb.jpeg"))

@@

# Methods

>>>>

In Figure \@ref(fig:simulation-intervals-state), we see the simulation intervals for all two-week intervals and all states as well as the 95% Covidestim credible intervals summed to be on the same time scale. 
We see agreement is much higher before the time period spanning December 2021 through January 2022, where Covidestim intervals tend to be higher than the probabilistic bias intervals. This period corresponds to the Omicron wave of the pandemic (Supplementary Figure michigan-variant). This was a major shift in the pandemic: the Omicron variant is highly transmissible compared to previous variants, and it has immune invasion capacity, which means vaccines provides less protection against infection [@andrews2022]. Omicron also is associated with a lower infection fatality ratio [@liu2022], as well as milder infection overall, which may influence testing behavior. Its rise to dominance was rapid, presumably as a result of this enhanced transmissibility and immune escape;  it became the dominant variant over the course of a single month.

@@

ADD TEST POSITIVITY FIGURE
FIGURE OUT WHICH STATES HAVE LAGS

@@

We focus on Massachussetts because this state is the only state that reports both positive and total tests at the county-level and has wastewater data consistently reported throughout the entire time period of interest. Wastewater data here is useful in the sense that it is inherently not subject to the same biases as voluntary testing data, since it captures anyone in the wastewater catchment area, and as such is not influenced by more symptomatic people being more likely to be tested or differences in access to testing.

@@

```{r testrate-v-ratio, fig.cap ="The ratio of the median estimated infections to observed infections plotted against the testing rate, where the testing rate is calculated as the total number tested in a two-week interval over the population size. When the priors are the same for all time intervals, there is minimal variability relationship between the testing rate and the ratio of estimated to observed infections, since the correction for incomplete testing and diagnostic test inaccuracy is identical for each time-interval. However, when we allow $\\beta$ or $\\Pr(S_1|\\text{untested})$ to vary over time, there is more variability in the relationship. A horizontal line in red at 1 is included to reference; a ratio of exactly 1 would indicate no infections went unobserved."}

@@

We see in Figure \@ref(fig:correlations) (a) that for all implementations, the probabilistic bias estimates are highly correlated with wastewater concentrations, and in Figure \@ref(fig:correlations) (b) we see they also highly correlated with the Covidestim estimates. The correlations in both cases are very similar between implementations, differing by less than 0.01.

@@

In Table \@ref(table:coverage), we summarize the concordance with Covidestim at the county and state levels by comparing the percent of simulation intervals where the Covidestim median falls above, below, or within the interval. At both geographic scales, we see that the probabilistic bias implementation that is most concordant with Covidestim is the that where the prior for $\Pr(S_1|\text{untested})$ is centered at the survey value of the percent of the population experiencing COVID-19-like illness, followed by the implementation that does not vary by state or date. This agreement was not a result of wider bounds, as the simulation intervals for the other two implementations were larger on average. For both of these implementations, when the Covidestim median was not contained, this was typically because it fell above the simulation interval. As is clear from Figure \@ref(fig:simulation-intervals-state) at the state-level, this often occurred during the Omicron wave, when the Covidestim estimates are consistently above the probabilistic bias counts.

@@

county_coverage %>%
  bind_rows(state_coverage) %>%
  group_by(`Geographic Scale`) %>%
  arrange(desc(`Percent Contained in Interval`),
          .by_group=TRUE) %>%
  mutate(across(contains("percent"),~ paste0(round(.,2), "\\%"))) %>%
  ungroup() %>%
  select(-`Geographic Scale`) %>%
  kbl(caption = "\\label{table:coverage}Coverage of Covidestim Medians at the County and State Levels",
      booktabs=TRUE,  
      escape=FALSE,
      digits=3) %>%
  kable_styling(latex_options=c("scale_down","HOLD_position"))  %>%
  group_rows(start_row=1, end_row =4, group_label= "County", 
             latex_gap_space = "0.7em",
             hline_after = TRUE) %>%
  group_rows(start_row=5, end_row =8, group_label= "State",
             latex_gap_space = "0.7em",
             hline_after = TRUE) %>%
  row_spec(1:(row_num-1),  hline_after = TRUE, 
           extra_css = "border-bottom: 1px solid;") %>%
  row_spec(0, font_size=11, bold=TRUE) %>%
  footnote(general = paste0("The percent of simulation intervals where the Covidestim median falls below, within, or above the interval,",
  "when considering all simulation intervals for that implementation and geographic scale. ",
  "Implementations are ordered from those with the highest proportion of Covidestim medians contained in the interval, to the lowest."),
  general_title ="", footnote_as_chunk=TRUE, threeparttable=TRUE)

@@

In Figure \@ref(fig:cross-corr-wastewater), we see that among all counties except one in Massachsuetts, the maximum correlation between mean wastewater concentration and biweekly infection estimates was obtained at a lag of zero, that is, there  typically there was no lag among the biweekly infection estimates and mean wastewater concentrations. This is to be expected given the lead time for wastewater concentrations, when observed, is less than 2 weeks [@olesen2021a]. In two counties, Hampsphire and Barnstable, the lag between the probabilistic bias counts and wastewater concentration was less than that between the observed positive tests and wastewater concentrations.

```{r cross-corr-wastewater, fig.width=8, fig.cap="Results of cross correlation analysis between probabilistic bias estimates and wastewater concentrations aggregated to the county level for counties in Massachusetts. For each county, the maximum correlation between the probablistic bias estimates and wastewater concentration is shown, colored by implementation; also included is the correlation between the observed positive tests and wastewater concentration. The shape indicates the lag at which the maximum correlation occurred, with units in biweeks. For example, a lag of -1 indicates the maximum correlation is obtained when the biweekly concentrations are lagged by 1 biweek, or equivalently, that the wastewater concentrations lead the observed positive tests by one biweek. For all counties except Barnstable, the maximum correlations between wastewater concentrations and probablistic bias estimates occur at no lag. In Barnstable and Hampshire, observed tests lagged wastewater concentrations, but the probablistic bias estimates for these counties did not."}

include_graphics(here("figures/cross_correlation_observed_pb.pdf"))

@@

# Methods {#Methods}
